```{r, include=FALSE}
species_name <- "Acanthagenys rufogularis"
```

```{r}
species_title(species_name, species_data)
```

Insert header text here.

```{r}
render_species_profile(species_name, caption = "Rare vagrant to the Brisbane region, with only a handful of records, all in winter months indicating seasonal movements. Common throughout the interior of Australia but rarely makes it to the coast.")
```

```{r}
render_species_table(species_name)
```

```{r, eval = isTRUE(knitr:::is_html_output())}
render_species_audio(species_name, audio_data)
```

```{r}
start_two_column_layout()
```

The Spiny-cheeked Honeyeater is a large, distinctive species of honeyeater similar (and closely related) to wattlebirds, but with a distinctive tawny breast and throat. It is gregarious in its normal range, although the majority of records in Brisbane have been on single birds or occasionally pairs. It is nomadic across much of its range, particularly in the north of Australia, although the provenance of all Brisbane birds is unknown. It is readily identified by its size (large compared to most honeyeaters), distinctive call and plumage pattern. Further birds may turn up anywhere across Brisbane, although there appears to be a preference for open woodland and dry forest, habitat similar to that found in their normal breeding range.

```{r}
end_two_column_layout()
```

```{r}
clear_page()
```

```{r, eval = isTRUE(knitr:::is_html_output())}
render_species_widget(species_name)
```

```{r, eval = !isTRUE(knitr:::is_html_output())}
render_species_map(species_name)
```

```{r}
render_species_graph(species_name)
```

```{r}
clear_page()
```

```{r}
start_two_column_layout()
```

In ut elit facilisis, laoreet nulla et, porta leo. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Phasellus ullamcorper eros at magna malesuada, ac maximus nibh imperdiet. Pellentesque id purus nec erat pharetra iaculis. Mauris velit neque, faucibus sed faucibus et, lobortis et felis. Morbi suscipit odio pretium pellentesque finibus. Aenean ultrices ex sapien. Curabitur tempus justo dolor, sit amet pulvinar diam fringilla at. Cras sit amet auctor ante. Maecenas laoreet nisi sem, elementum condimentum sem aliquam in. Praesent aliquam consequat leo sed fringilla. Aliquam sit amet metus quis ligula ultricies pharetra ut vitae velit. Sed sed vulputate mi. Nullam et elit vitae arcu efficitur venenatis.

In ut elit facilisis, laoreet nulla et, porta leo. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Phasellus ullamcorper eros at magna malesuada, ac maximus nibh imperdiet. Pellentesque id purus nec erat pharetra iaculis. Mauris velit neque, faucibus sed faucibus et, lobortis et felis. Morbi suscipit odio pretium pellentesque finibus. Aenean ultrices ex sapien. Curabitur tempus justo dolor, sit amet pulvinar diam fringilla at. Cras sit amet auctor ante. Maecenas laoreet nisi sem, elementum condimentum sem aliquam in. Praesent aliquam consequat leo sed fringilla. Aliquam sit amet metus quis ligula ultricies pharetra ut vitae velit. Sed sed vulputate mi. Nullam et elit vitae arcu efficitur venenatis.

In ut elit facilisis, laoreet nulla et, porta leo. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Phasellus ullamcorper eros at magna malesuada, ac maximus nibh imperdiet. Pellentesque id purus nec erat pharetra iaculis. Mauris velit neque, faucibus sed faucibus et, lobortis et felis. Morbi suscipit odio pretium pellentesque finibus. Aenean ultrices ex sapien. Curabitur tempus justo dolor, sit amet pulvinar diam fringilla at. Cras sit amet auctor ante. Maecenas laoreet nisi sem, elementum condimentum sem aliquam in. Praesent aliquam consequat leo sed fringilla. Aliquam sit amet metus quis ligula ultricies pharetra ut vitae velit. Sed sed vulputate mi. Nullam et elit vitae arcu efficitur venenatis.

In ut elit facilisis, laoreet nulla et, porta leo. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Phasellus ullamcorper eros at magna malesuada, ac maximus nibh imperdiet. Pellentesque id purus nec erat pharetra iaculis. Mauris velit neque, faucibus sed faucibus et, lobortis et felis. Morbi suscipit odio pretium pellentesque finibus. Aenean ultrices ex sapien. Curabitur tempus justo dolor, sit amet pulvinar diam fringilla at. Cras sit amet auctor ante. Maecenas laoreet nisi sem, elementum condimentum sem aliquam in. Praesent aliquam consequat leo sed fringilla. Aliquam sit amet metus quis ligula ultricies pharetra ut vitae velit. Sed sed vulputate mi. Nullam et elit vitae arcu efficitur venenatis.

```{r}
end_two_column_layout()
```

```{r}
clear_page()
```
