```{r, include=FALSE}
species_name <- "Acanthiza chrysorrhoa"
```

```{r}
species_title(species_name, species_data)
```

Insert header text here.

```{r}
render_species_profile(species_name, caption = "Patchy resident within the Brisbane region with only a handful of regular sites but many more locations where the species has occurred. Present all year round with a minor increase over winter months, and no clear trend over the past 10 years. Not of conservation concern within Brisbane or in wider Australia, as the species is common throughout the majority of the southern states in suitable habitat.")
```

```{r}
render_species_table(species_name)
```

```{r, eval = isTRUE(knitr:::is_html_output())}
render_species_audio(species_name, audio_data)
```

```{r}
start_two_column_layout()
```

The Yellow-rumped Thornbill is a distinctive small passerine with a bright yellow rump and conspicuous facial markings. It is a member of the family Acanthizidae, which includes other thornbills, gerygones and scrubwrens. It is the largest species of thornbill, up to 12cm in length [@HANZAB6] and diverged early from other thornbills [@marki2017supermatrix], resulting in a species quite distinct from the rest within its genus. If seen well, it can be identified easily and calls are further able to differentiate the birds from other species within the Acanthizidae. Within Brisbane, Yellow-rumped Thornbills are patchily distributed and are only common at a few sites, but they wander widely and can turn up anywhere with suitable habitat. Most commonly found in small groups of 2-4, numbers of up to a dozen have been recorded rarely.

```{r}
end_two_column_layout()
```

```{r}
clear_page()
```

```{r, eval = isTRUE(knitr:::is_html_output())}
render_species_widget(species_name)
```

```{r, eval = !isTRUE(knitr:::is_html_output())}
render_species_map(species_name)
```

```{r}
render_species_graph(species_name)
```

```{r}
clear_page()
```

```{r}
start_two_column_layout()
```

A species of dry woodlands and open forests, the Yellow-rumped Thornbill is well-adapted to life in Australian environments, and despite its patchy occurrence within Brisbane the species is stable in the region. Common across much of Western Australia, South Australia, the southern Northern Territory and the south east States, these birds are quintessential Australian bush inhabitants.

Yellow-rumped Thornbills have been recorded from all across mainland Brisbane excluding the wetter forests of the Camel's Head and the south east corner (and the species has never been recorded on Moreton Island), but they are only patchily common and occur in small numbers. Strongholds are at Prior's Pocket, Oxley Creek Common and the grassy woodlands near the mouth of the Brisbane River, where the species occurs widely. The highest count is of 12 birds at Oxley Creek Common [@possingham2009ebirdS17747770] but typical counts are much lower, with only a handful of records over 5 birds.

The species is present in Brisbane year-round, although there is a slight increase in reporting rate over the winter months, indicating a degree of seasonality.

**Distribution and Habitat**

Yellow-rumped Thornbills are found across Brisbane, with a preference for open habitats at lower altitudes, although the species has been recorded up to 300m. Most commonly found in non-remnant and estuarine habitats, the birds favour open land with some tree cover to hide in. Given the abundance of this habitat within Brisbane, and the commonness of the species across much of Australia, it is somewhat surprising that Yellow-rumped Thornbills are as uncommon as they are within the LGA. Even at the highest rate of reporting in June, the species is only found on about 2.5% of checklists, so while the birds are found quite widely across Brisbane, they are by no means common.

**Seasonality**

The species is present in Brisbane all year, although there is an increase in records over the cooler months, as noted above. Breeding is scarcely reported, with a handful of records in June, September and October. It would be good to collect more breeding data on this species, and to find out more about its seasonal movements and any evidence of nomadism.

**Trends**

The reporting rate for Yellow-rumped Thornbills has been very variable between 2005 and 2017, with bumper years in 2008 and 2011 (reported on ~5% of checklists) and other years such as 2014, 2016 and 2017 being much lower (reported on ~1% of checklists). More work needs to be done to investigate the reasons for such fluctuation, but no clear trend is apparent from the data. Outside of Brisbane, the species is much more common across a large area of the continent and as such is of no conservation concern.

**Information Gaps**

- Identify the reasons behind any erratic reporting rates across years
- Identify causes behind seasonality and breeding patterns
- Increased detail in reporting, including audio, photos and breeding data

**Key Conservation Needs**

- Protect favourable habitats for this species and identify any local declines


```{r}
end_two_column_layout()
```

```{r}
clear_page()
```
