```{r, include=FALSE}
species_name <- "Acanthiza lineata"
```

```{r}
species_title(species_name, species_data)
```

Insert header text here.

```{r}
render_species_profile(species_name, caption = "Limited to the wetter rainforests of the inland ranges of Brisbane, a patchy resident and possible partial migrant. Slightly more records in cooler months, but no clear indication of local movements. Occurrence varies widely from year to year with no clear long term trend. Not of conservation concern, but dependent on the cooler wet sclerophyll and rainforests of the Brisbane ranges, which are under threat.")
```

```{r}
render_species_table(species_name)
```

```{r, eval = isTRUE(knitr:::is_html_output())}
render_species_audio(species_name, audio_data)
```

```{r}
start_two_column_layout()
```

The Striated Thornbill is a small, brightly-coloured thornbill with a distinctively streaked face and underparts. It is a member of the family Acanthizidae alongside other thornbills, gerygones and scrubwrens, and is in the genus Acanthiza like all other thornbill species. A gregarious species, it is found in pairs or small groups, with flocks of up to half a dozen being moderately common in the Brisbane region. Identification can be somewhat challenging without clear views, but their distinctive calls and brightly coloured bodies separate them from most other thornbill species in the area. Within Brisbane, Striated Thornbills are patchily distributed, occurring commonly at a number of sites with suitable wet forest habitat, but are absent at others. Given that the majority of the habitat within Brisbane in which they may occur is poorly surveyed, they may be more common than currently indicated.

```{r}
end_two_column_layout()
```

```{r}
clear_page()
```

```{r, eval = isTRUE(knitr:::is_html_output())}
render_species_widget(species_name)
```

```{r, eval = !isTRUE(knitr:::is_html_output())}
render_species_map(species_name)
```

```{r}
render_species_graph(species_name)
```

```{r}
clear_page()
```

```{r}
start_two_column_layout()
```

In ut elit facilisis, laoreet nulla et, porta leo. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Phasellus ullamcorper eros at magna malesuada, ac maximus nibh imperdiet. Pellentesque id purus nec erat pharetra iaculis. Mauris velit neque, faucibus sed faucibus et, lobortis et felis. Morbi suscipit odio pretium pellentesque finibus. Aenean ultrices ex sapien. Curabitur tempus justo dolor, sit amet pulvinar diam fringilla at. Cras sit amet auctor ante. Maecenas laoreet nisi sem, elementum condimentum sem aliquam in. Praesent aliquam consequat leo sed fringilla. Aliquam sit amet metus quis ligula ultricies pharetra ut vitae velit. Sed sed vulputate mi. Nullam et elit vitae arcu efficitur venenatis.

In ut elit facilisis, laoreet nulla et, porta leo. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Phasellus ullamcorper eros at magna malesuada, ac maximus nibh imperdiet. Pellentesque id purus nec erat pharetra iaculis. Mauris velit neque, faucibus sed faucibus et, lobortis et felis. Morbi suscipit odio pretium pellentesque finibus. Aenean ultrices ex sapien. Curabitur tempus justo dolor, sit amet pulvinar diam fringilla at. Cras sit amet auctor ante. Maecenas laoreet nisi sem, elementum condimentum sem aliquam in. Praesent aliquam consequat leo sed fringilla. Aliquam sit amet metus quis ligula ultricies pharetra ut vitae velit. Sed sed vulputate mi. Nullam et elit vitae arcu efficitur venenatis.

In ut elit facilisis, laoreet nulla et, porta leo. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Phasellus ullamcorper eros at magna malesuada, ac maximus nibh imperdiet. Pellentesque id purus nec erat pharetra iaculis. Mauris velit neque, faucibus sed faucibus et, lobortis et felis. Morbi suscipit odio pretium pellentesque finibus. Aenean ultrices ex sapien. Curabitur tempus justo dolor, sit amet pulvinar diam fringilla at. Cras sit amet auctor ante. Maecenas laoreet nisi sem, elementum condimentum sem aliquam in. Praesent aliquam consequat leo sed fringilla. Aliquam sit amet metus quis ligula ultricies pharetra ut vitae velit. Sed sed vulputate mi. Nullam et elit vitae arcu efficitur venenatis.

In ut elit facilisis, laoreet nulla et, porta leo. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Phasellus ullamcorper eros at magna malesuada, ac maximus nibh imperdiet. Pellentesque id purus nec erat pharetra iaculis. Mauris velit neque, faucibus sed faucibus et, lobortis et felis. Morbi suscipit odio pretium pellentesque finibus. Aenean ultrices ex sapien. Curabitur tempus justo dolor, sit amet pulvinar diam fringilla at. Cras sit amet auctor ante. Maecenas laoreet nisi sem, elementum condimentum sem aliquam in. Praesent aliquam consequat leo sed fringilla. Aliquam sit amet metus quis ligula ultricies pharetra ut vitae velit. Sed sed vulputate mi. Nullam et elit vitae arcu efficitur venenatis.

```{r}
end_two_column_layout()
```

```{r}
clear_page()
```
