```{r, include=FALSE}
species_name <- "Acridotheres tristis"
```

```{r}
species_title(species_name, species_data)
```

Insert header text here.

```{r}
render_species_profile(species_name, caption = "Widespread introduced pest species that appears to be declining within Brisbane despite rapid expansion across much of Australia. Appears to exhibit a degree of seasonal variation within Brisbane, being nearly twice as abundant over summer, but no clear reason for this pattern. Not of conservation concern owing to pest status.")
```

```{r}
render_species_table(species_name)
```

```{r, eval = isTRUE(knitr:::is_html_output())}
render_species_audio(species_name, audio_data)
```

```{r}
start_two_column_layout()
```

The Common Myna is an introduced species of Starling, native to south east Asia and the Indian Subcontinent. It is unmistakeable within the Brisbane region, with brown and black plumage and a striking yellow beak and bare eye skin. Mynas are common around the built-up areas of the city, being less common in densely vegetated areas with remnant habitat. Data from the past 15 years suggests this species is declining in Brisbane, with annual abundance dropping from 40% in 2005 to 20% in 2017, but reasons for this are unclear.

```{r}
end_two_column_layout()
```

```{r}
clear_page()
```

```{r, eval = isTRUE(knitr:::is_html_output())}
render_species_widget(species_name)
```

```{r, eval = !isTRUE(knitr:::is_html_output())}
render_species_map(species_name)
```

```{r}
render_species_graph(species_name)
```

```{r}
clear_page()
```

```{r}
start_two_column_layout()
```

A colonial species of open habitats around Brisbane, Common Mynas are an introduced pest species that are widely distributed across the region. Interestingly, although the species can be regularly found in large flocks, birds appear to be declining in abundance quite rapidly, with a 50% decrease in reporting rate over the last decade. Mynas are found in nearly any open areas of the city, especially in non-remnant, built-up, wetland and estuarine habitats.

Although the species is often found in large flocks of over 50 birds, the average count is much lower, sitting around 3.5 birds per checklist. The high count is 500 birds at Stones Corner, 

```{r}
end_two_column_layout()
```

```{r}
clear_page()
```
