```{r, include=FALSE}
species_name <- "Aerodramus terraereginae"
```

```{r}
species_title(species_name, species_data)
```

Category Pending; Needs verification.

```{r}
render_species_profile(species_name, caption = "Endemic swiftlet of north east Queensland. One report from Brisbane region but no documentation, unverified. Not of conservation concern, but range somewhat restricted.")
```

```{r}
render_species_table(species_name)
```

```{r, eval = isTRUE(knitr:::is_html_output())}
render_species_audio(species_name, audio_data)
```

```{r}
start_two_column_layout()
```

The Australian Swiftlet is a medium-sized swiftlet endemic to coastal north east Queensland where it breeds in caves and gorges. It is the only regular species of swiftlet in Australia and is readily separable with decent views from needletails and swifts. A sedentary species, there are very few confirmed records south of the edge of this species' breeding range in Mackay, and all reports to date have been unverified.

```{r}
end_two_column_layout()
```

```{r}
clear_page()
```

```{r, eval = isTRUE(knitr:::is_html_output())}
render_species_widget(species_name)
```

```{r, eval = !isTRUE(knitr:::is_html_output())}
render_species_map(species_name)
```

```{r}
render_species_graph(species_name)
```

```{r}
clear_page()
```

```{r}
start_two_column_layout()
```

A colonial species of savannas and open woodland, Australian Swiftlets can reach high abundances within their breeding range, but are quite sedentary and rarely move beyond their favoured locations. Within Brisbane, there is one record of the species, from 2014 at Sandy Camp Road [@ringma2014ebirdS16784764], indicating that at best they are a very rare vagrant.

**Distribution and Habitat**

Australian Swiftlets are typically found in north east Queensland from the Iron Range to Mackay and inland to Chillagoe (where there is a separate subspecies). They favour open savanna woodlands mainly, but can be found in a variety of habitats at most lowland elevations. They are sedentary within their home range and there are very few vagrant records on eBird, indicating that despite being highly mobile and adept fliers, birds do not tend to move very far to feed. The closest records from outside Brisbane are in the Lockyer Valley and south of Tweed Heads.

**Seasonality and Breeding**

Further birds could conceivably arrive in Brisbane at any time of the year, as the species are not migratory nor do they exhibit any seasonal chances in abundance within their normal range. Extreme weather events, such as tropical cyclones or extreme low fronts may play a role in pushing birds south, so it is worth checking all swift and needletail flocks for this species (or other vagrants) following such events.

**Trends**

Australian Swiftlets are not currently threatened within their home range.

**Information Gaps**

- Verify current eBird record and locate any historical records not currently in the eBird database
- Identify any causes of vagrant behaviour

**Key Conservation Needs**

- Protect breeding and feeding habitats in normal range

**Contributors to Species Account**

Louis Backstrom


```{r}
end_two_column_layout()
```

```{r}
clear_page()
```
