```{r, include=FALSE}
species_name <- "Amaurornis moluccana"
```

```{r}
species_title(species_name, species_data)
```

Category A; Uncommon localised resident.

```{r}
render_species_profile(species_name, caption = "Rare and secretive resident of riparian habitats around the city. Seasonal variations in reporting rate most likely a result of this species' conspicuous calls during the breeding season but may also be indicative of annual movements.")
```

```{r}
render_species_table(species_name)
```

```{r, eval = isTRUE(knitr:::is_html_output())}
render_species_audio(species_name, audio_data)
```

```{r}
start_two_column_layout()
```

A drab-plumaged bird, superficially similar to the closely-related but much more common [Dusky Moorhen](#dusky-moorhen-gallinula-tenebrosa), the Pale-vented Bush-hen is a secretive species of rail found across much of the city wherever habitat is suitable. Birds are typically found individually or in pairs, although small family groups have been recorded at times.

```{r}
end_two_column_layout()
```

```{r}
clear_page()
```

```{r, eval = isTRUE(knitr:::is_html_output())}
render_species_widget(species_name)
```

```{r, eval = !isTRUE(knitr:::is_html_output())}
render_species_map(species_name)
```

```{r}
render_species_graph(species_name)
```

```{r}
clear_page()
```

```{r}
start_two_column_layout()
```

A skulking resident of dense vegetation lining creeks and gullies around the city, the Pale-vented Bush-hen is one of Brisbane's rarer and more difficult to find species of waterbird, although birds can generally be located with some persistence at the right sites. Birds have been reported along the verges of most of Brisbane's major creek systems, including Cabbage Tree Creek, Kedron Brook, Enoggera Creek, Oxley Creek and Bulimba Creek. Birds also occasionally turn up in other locations, but are generally found not too far away from water and suitable lining vegetation.

Birds are present in Brisbane year-round, although are reported much more frequently during the warmer months of the year between September and March. This is likely a result of birds being more active, showy and vocal during the breeding season in spring and summer, but may also be indicative of actual movements; this requires further investigation. Birds appear to be quite well-adapted to living in human-modified habitats, although at the same time seem to be quite wary of human presence, so are at moderate risk of disturbance and declines should they lose habitat in the future. Nevertheless, their population appears to currently be stable.

**Distribution and Habitat**



```{r}
end_two_column_layout()
```

```{r}
clear_page()
```
