```{r, include=FALSE}
species_name <- "Anous minutus"
```

```{r}
species_title(species_name, species_data)
```

Category A; Rare widespread summer migrant.

```{r}
render_species_profile(species_name, caption = "Rare visitor to offshore waters during the summer months from the north, may be more common than current information suggests. Not of any conservation concern at present.")
```

```{r}
render_species_table(species_name)
```

```{r, eval = isTRUE(knitr:::is_html_output())}
render_species_audio(species_name, audio_data)
```

```{r}
start_two_column_layout()
```

The Black Noddy is the rarer of the two main Noddy species found in Brisbane, and can be found offshore throughout the summer months in low numbers. It is smaller and darker than its close relative the [Common Noddy](#common-noddy-anous-stolidus), with a long bill a contrasting white cap. Birds typically stay well offshore but do come in close, even within the bay, in inclement weather.

```{r}
end_two_column_layout()
```

```{r}
clear_page()
```

```{r, eval = isTRUE(knitr:::is_html_output())}
render_species_widget(species_name)
```

```{r, eval = !isTRUE(knitr:::is_html_output())}
render_species_map(species_name)
```

```{r}
render_species_graph(species_name)
```

```{r}
clear_page()
```

```{r}
start_two_column_layout()
```

Widely distributed across the Pacific Ocean and parts of the Central Atlantic and Indian Oceans, the Black Noddy is a small, slender species of tern found in tropical and subtropical waters. Within Australia, the species breeds on islands off the north half of the continent as far south as Lady Elliot Island (with further colonies on Lord Howe and Norfolk Islands as well) and rarely comes close to land, although notable records of this species have occurred from within Moreton Bay following cyclonic conditions out to sea.

Birds could turn up in Brisbane waters year-round but are probably most common off the coast in summer and autumn; this is when the species has been regularly recorded off south-east Queensland pelagic trips. Very little current data exists on the occurrence of this species (and indeed most seabirds) off the coast of Moreton Bay, and without more information it is difficult to predict with certainty when birds are most common in the city. Several threats, primarily nest predation from introduced predators on breeding islands, imperil this species across its range but at present populations are generally stable and the species is not considered endangered.

**Distribution and Habitat**

Within Brisbane, Black Noddies are almost exclusively pelagic, typically being found off the coast out near the continental shelf break. Birds do occasionally venture in closer to shore and may be seen while seawatching, particularly during periods of inclement weather such as following tropical cyclones. Their feeding pattern does not appear to be particularly dependent on seasonal changes, although finer-scale monitoring is likely needed to confirm this.

**Seasonality**

As noted above, Black Noddies do not breed in Brisbane; the closest breeding colonies to the city are Lady Ellio Island, 400km to the north, and Lord Howe Island, 750km to the south east. Birds are present around these colonies throughout most of the year, although they are less common during the winter months. Within Brisbane they appear to be most abundant during late summer and early autumn, although more data are needed to verify this.

**Trends**

Once again, due to the paucity of data available for this species in south east Queensland, it is difficult to draw any firm conclusions about this species' local status in the region. However, the species is not globally or nationally threatened and most populations appear to be stable. Key threats include predation of nests on breeding islands, particularly by introduced mammalian predators, and rising sea levels as a result of climate change.

**Information Gaps**

- Collect more records of this species off the coast

**Key Conservation Needs**

- Monitor populations at nearby breeding locations
- Monitor frequency of local records for any changes

**Contributors to Species Account**

Louis Backstrom

```{r}
end_two_column_layout()
```

```{r}
clear_page()
```
