---
title: "Atlas of the Birds of Brisbane"
date: "`r Sys.Date()`"
documentclass: book
author: "Created by the birdwatching community of Brisbane."
classoption:
  - openany
  - oneside
fontsize: 12pt
bibliography: references.bib
csl: reference-style.csl
biblio-style: reference-style
link-citations: yes
github-repo: bird-team/brisbane-bird-atlas
favicon: "assets/misc/favicon.ico"
url: "https\\://brisbanebirds.com/"
description: "Community bird atlas for Brisbane, Australia"
geometry: "left=2.54cm, right=2.54cm, top=2.54cm, bottom=2.54cm"
---

```{r, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE,
                      results = "asis", fig.align = "center")
knitr::opts_knit$set(progress = FALSE, verbose = FALSE)
```

```{r, include = FALSE}
source("code/scripts/index.R")
```

```{r}
start_page()
```
# Welcome!

The _Atlas_ provides information on the distribution, abundance, breeding, seasonality and trends for every bird species known to have occurred within the Brisbane Local Government Area boundary, and the adjacent coastal waters and Coral Sea. Brisbane is one of the most biodiverse cities in the world, and the Atlas is a way to keep track of its avian inhabitants, ensuring they continue to have space to thrive alongside our bustling human communities. This project officially launches on 1 January 2019, at which point it will be open for public participation. The goal is to draw on the collective expertise of birdwatchers throughout the city, and those who have visited from across the world, and record in one place all the information we have about Brisbane's birds. The _Atlas_ is a living, online, open-access resource that will continue to grow through community input as the years go by. The plan is to have the first full draft of the _Atlas_ completed by 2020, and then continue to produce annual editions thereafter. The _Atlas_ is downloadable in PDF form (click the link at the top of this page), and is accessible by browsing this website.

We hope you enjoy looking through the content - check back frequently as the _Atlas_ grows. If you can, please also donate your time and expertise to help make the _Atlas_ the best it can be. For full details of how you can contribute, please see the [Contributing](https://brisbanebirds.com/contributing.html) section. Here are some of the key things you can do to help create this landmark resource:

1. Go birdwatching in the city of Brisbane (see [here](http://www.dlgrma.qld.gov.au/resources/map/reform/brisbane-map.pdf) for a map) and enter your observations on [eBird](http://www.ebird.org), a free online tool for recording bird observations. All eBird records for Brisbane will be automatically included in the _Atlas_.

2. If you have old records stored away on notebooks and such, enter them into eBird too! See [here](https://help.ebird.org/customer/en/portal/articles/973960-entering-historic-data) on more information about entering historic data into eBird. If you want any help with importing your records, email [Richard Fuller](mailto:r.fuller@uq.edu.au) to discuss - we will be glad to help you, or even to fully manage the task on your behalf and get you set up on eBird.

3. If you have taken any better images than those already used as the header photo in a species account, please consider contributing them to the _Atlas_. The photo must have been taken in Brisbane, and should be cropped to 16:9. The image will be licensed in the public domain, hosted on the [Atlas Flickr site](https://www.flickr.com/photos/159495711@N04/), and always used with proper attribution to the photographer in the _Atlas_ - email it to [Richard Fuller](mailto:r.fuller@uq.edu.au).

4. Write some text for the _Atlas_, or edit and improve some of the existing text. See [here](https://github.com/bird-team/brisbane-bird-atlas/wiki/Frequently-asked-questions#how-can-i-contribute-to-this-atlas) for details. Full instructions on how to edit _Atlas_ pages will be available in the [Contributing](https://brisbanebirds.com/contributing.html) section from 1 January 2019 onwards.

5. If you have any suggestions for content, or information about Brisbane's birds that you think would be helpful for the _Atlas_, please contact the editors. Email [Richard Fuller](mailto:r.fuller@uq.edu.au), or create an issue on the [GitHub Repository](https://github.com/bird-team/brisbane-bird-atlas/issues).

Above all, enjoy Brisbane's birds!

_Book Cover and Website Icon design by Louis Backstrom with images sourced from Wikimedia Commons under Creative Commons Attribution._
