---
title: "Atlas of the Birds of Brisbane"
date: "`r Sys.Date()`"
documentclass: book
author: "Created by the birdwatching community of Brisbane."
classoption:
  - openany
  - oneside
fontsize: 12pt
bibliography: references.bib
csl: reference-style.csl
biblio-style: reference-style
link-citations: yes
github-repo: bird-team/brisbane-bird-atlas
favicon: "assets/misc/favicon.ico"
url: "https\\://brisbanebirds.com/"
description: "Community bird atlas for Brisbane, Australia"
geometry: "left=2.54cm, right=2.54cm, top=2.54cm, bottom=2.54cm"
---

```{r, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE,
                      results = "asis", fig.align = "center")
knitr::opts_knit$set(progress = FALSE, verbose = FALSE)
```

```{r, include = FALSE}
source("code/scripts/index.R")
```

```{r}
start_page()
```
# Welcome!

The _Atlas_ provides information on the distribution, abundance, breeding, seasonality and trends for every bird species known to have occurred within the Brisbane Local Government Area boundary, and the adjacent coastal waters and Coral Sea. Brisbane is one of the most biodiverse cities in the world, and the _Atlas_ is a way to keep track of its avian inhabitants, ensuring they continue to have space to thrive alongside our bustling human communities.

This project officially launches on 1 January 2019, at which point it will be open for public participation. The goal is to draw on the collective expertise of birdwatchers throughout the city, and those who have visited from across the world, recording in one place all the information we have about Brisbane's birds. The _Atlas_ is a living, online, open-access resource that will continue to grow through community input as the years go by. The plan is to have the first full draft of the _Atlas_ completed by 2020, and then continue to produce annual editions thereafter. The _Atlas_ is downloadable in PDF form (click the link at the top of this page), and is accessible by browsing this website.

We hope you enjoy looking through the content - check back frequently as the _Atlas_ grows. If you can, please also donate your time and expertise to help make the _Atlas_ the best it can be. For full details of how you can contribute, please see the [Contributing](https://brisbanebirds.com/contributing.html) section.

Above all else, enjoy Brisbane's birds!

_Book Cover and Website Icon design by Louis Backstrom with images sourced from Wikimedia Commons under Creative Commons Attribution._
